---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# noon

<!-- badges: start -->
<!-- badges: end -->

`{noon}` allows to watch for incoming MIDI events from R. 
Because why not. 

## Install

You'll need a running installation of NodeJS, that you can get from [nodejs.org](https://nodejs.org/en/download/). 

Then, the package can be installed with:

```{r}
remotes::install_github("ColinFay/noon")
```

After install, you'll need to run __one time__ `noon_npm_install()` to set up all the Node dependencies.
This will ask you if you're OK with doing that installation.

```{r}
library(noon)
noon_npm_install()
```

## Functions reference

## `midi_list()`

List all the available midi interfaces:

```{r}
midi_list()
```

### `midi_all()` 

`midi_all()` will launch a process that checks every `n` seconds (default 0.5) if a new midi event has been received on all the midi interface. 

It takes a function defining what to do with the message (note it's received in JSON). 
By default, R will `cat()` the output. 

```{r}
midi_all()
```

This function takes the following arguments: 

+ callback: The callback function to execute when R receives a MIDI message. 
This function takes one argument, which is the JSON MIDI message.
+ n: time in second between each check for a new MIDI message. 
+ with_clock: whether or not to listen to the clock event.

The message is received with the following format: 

```json
{"input":"microKONTROL PORT A","event":"cc","msg":{"channel":4,"controller":13,"value":127,"_type":"cc"}}
```


### `midi_close_all()` 

Don't forget to `midi_close_all()`, if you don't want to watch on MIDI inputs anymore.

```{r}
midi_close_all()
```

### `midi_one_event()`

This function will listen to one event type only (all interfaces)

```{r}
midi_one_event(event = "cc")
```

This function takes the following arguments: 

+ callback: The callback function to execute when R receives a MIDI message. 
This function takes one argument, which is the JSON MIDI message.
+ n: time in second between each check for a new MIDI message. 
+ event: the event to listen to. Available event are "noteoff", "noteon", "poly aftertouch", "cc", "program", "channel aftertouch", "pitch", "position", "select", "start", "continue", "stop", and "reset"

## `midi_one_interface()`

This function will listen to one interface type only (all events)

```{r}
midi_one_interface(interface = "microKONTROL PORT A")
```

This function takes the following arguments: 

+ interface: name of the interface to listen to. 
Must be one of `midi_list()`.
+ callback: The callback function to execute when R receives a MIDI message. 
This function takes one argument, which is the JSON MIDI message.
+ n: time in second between each check for a new MIDI message. 
+ with_clock: whether or not to listen to the clock event.

## `midi_one_event_one_interface()`

This function will listen to one event on one interface

```{r}
midi_one_event_one_interface(event = "cc", interface = "microKONTROL PORT A")
```

+ event: the event to listen to. Available event are "noteoff", "noteon", "poly aftertouch", "cc", "program", "channel aftertouch", "pitch", "position", "select", "start", "continue", "stop", and "reset"
+ interface: name of the interface to listen to. 
Must be one of `midi_list()`.
+ callback: The callback function to execute when R receives a MIDI message. 
This function takes one argument, which is the JSON MIDI message.
+ n: time in second between each check for a new MIDI message. 
+ with_clock: whether or not to listen to the clock event.

## About the `callback` argument

Each function takes a `callback` parameter, which define the logic to handle the message (received as a JSON string) under the following format: 

```json
{"input":"microKONTROL PORT A","event":"cc","msg":{"channel":4,"controller":13,"value":127,"_type":"cc"}}
```

> Important note: given that the functions checks the for MIDI midi on a recursive basis (i.e every `n` seconds), it's possible that at some point R receives more than one JSON strings in a row, for example if you 'noteon' and 'noteoff' very quickly. The `callback` will then be `lapply`-ed on all the inputs.

```{r}
handle <- function(x){
  print(jsonlite::fromJSON(x))
}
midi_all(callback = handle)
```

Note that it can also be written as a formula, following `{rlang}` syntax.

```{r}
midi_all(callback = ~ print(jsonlite::fromJSON(.x)))
```


### Some `callback` example

+ `load_all()` on a button press

```{r}
load_all <- function(x){
  x <- jsonlite::fromJSON(x)
  # listen on event 16 and only on press (value 127)
  if (x$msg$controller == 16 && x$msg$value == 127){
    cli::cat_line("Reloading the package")
    pkgload::load_all()
  }
}
midi_all(callback = load_all)
```

+ `ggplot()` on a button press

```{r}
library(ggplot2)
ggplotit <- function(x){
  x <- jsonlite::fromJSON(x)
  if (x$msg$controller == 15 && x$msg$value == 127){
    gg <- ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) + 
      geom_point()
    print(gg)
  }
}
midi_all(callback = ggplotit)
```

+ A dev oriented workflow

```{r}
handle <- function(x){
  x <- jsonlite::fromJSON(x)
  
  if (x$msg$controller == 11 && x$msg$value == 127){
    cli::cat_line("Opening README")
    rstudioapi::navigateToFile("README.Rmd")
  }
  
  if (x$msg$controller == 12 && x$msg$value == 127){
    cli::cat_line("Opening DESCRIPTION")
    rstudioapi::navigateToFile("DESCRIPTION")
  }
  
  if (x$msg$controller == 13 && x$msg$value == 127){
    cli::cat_line("Testing")
    devtools::test()
  }
  if (x$msg$controller == 14 && x$msg$value == 127){
    cli::cat_line("Checking")
    devtools::check()
  }
  if (x$msg$controller == 15 && x$msg$value == 127){
    cli::cat_line("Document")
    devtools::document()
  }
  if (x$msg$controller == 16 && x$msg$value == 127){
    cli::cat_line("Reloading the package")
    pkgload::load_all()
  }
}
midi_all(callback = handle)
```

+ `{beepr}`

```{r}
library(beepr)
handle <- function(x){
  x <- jsonlite::fromJSON(x)
  cli::cat_rule(x$msg$note)
  note <- switch(
    as.character(x$msg$note), 
    "60" = 1,
    "61" = 2,
    "62" = 3,
    "63" = 4,
    "64" = 5,
    "65" = 6,
    "66" = 7,
    "67" = 8,
    "68" = 9,
    "69" = 10,
    "70" = 11
  )
  beep(note)
  
}
midi_one_event(event = "noteon", callback = handle,  n = 0)
```

```{r include = FALSE}
library(audio)
handle <- function(x){
  x <- jsonlite::fromJSON(x)
  cli::cat_rule(x$msg$note)
  note <- switch(
    as.character(x$msg$note), 
    "60" = system.file("samples/C5.wav", package = "noon"),
    "61" = system.file("samples/C#5.wav", package = "noon"),
    "62" = system.file("samples/D5.wav", package = "noon"),
    "63" = system.file("samples/D#5.wav", package = "noon"),
    "64" = system.file("samples/E5.wav", package = "noon"),
    "65" = system.file("samples/F5.wav", package = "noon"),
    "66" = system.file("samples/F#5.wav", package = "noon"),
    "67" = system.file("samples/G5.wav", package = "noon"),
    "68" = system.file("samples/G#5.wav", package = "noon"),
    "69" = system.file("samples/A5.wav", package = "noon"),
    "70" = system.file("samples/A#5.wav", package = "noon"),
    "71" = system.file("samples/B5.wav", package = "noon")
  )
  if (file.exists(note)){
    withr::with_environment(
      .GlobalEnv, {
        x <- load.wave(note)
    play(x, rate = 44100 / 2)  
      }
    )
    
  }
  
}
midi_one_event(event = "noteon", callback = handle,  n = 0)
```
